---
title: "Practical Machine Learning Course Project"
author: "YHu"
date: "April 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Summary

The goal of this course project is to predict the manner of exericse -"classe" variable in the data, using other predictors in the dataset. I utilized random forest to do the prediction and validated on the test cases.

## Data Preparation
After reading the training and testing datasets, I took a look at the distribution of "classe" variable in the data.
```{r results='hide', message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(corrplot)
library(dplyr)
```

```{r, eval=TRUE}
train <-read.csv("pml-training.csv",header=TRUE, na.strings=c("NA",""))
test <- read.csv("pml-testing.csv",header=TRUE, na.strings=c("NA",""))
```

```{r echo=TRUE}
table(train$classe)
```

In addition, I also briefly viewed the data using the `head` function.
```{r, eval=FALSE}
head(train)
```

Noticing that there are a lot of missing values in the several variables, I removed from the data columns with all NAs. 

```{r echo=TRUE}
na_count <-function (x) sapply(x, function(y) sum(is.na(y)))
na_train<-(na_count(train)==0)
train <- train[,na_train]
```

Also columns that seem to be non-predictive were removed.
```{r echo=TRUE}
col.del <- grepl("X|user|timestamp|newwindow",colnames(train))
train <- train[,!col.del]
```

The training dataset was split into 70% train and 30% validation sets.
```{r echo=TRUE}
train.id<-createDataPartition(y=train$classe,p=0.7,list=FALSE)
train_new <- train[train.id,]
valid <- train[-train.id,]
```

I made the formula for the model to prepare for following model fitting.
```{r echo=TRUE}
predictors<-colnames(train_new%>%dplyr::select(-classe))
full.formula<-reformulate(predictors,"classe")
full.formula
```

##Fit the Random Forest model
A Random Forest model was fitted using all the predictors in the training dataset with "classe" as target.
```{r, eval=FALSE}
rf<-randomForest(full.formula,
                 data=train_new,
                 importance=TRUE,
                 ntree=500)
```
```{r echo=FALSE}
load("project.rda")
```
The model was then applied to the validation dataset.
```{r echo=TRUE}
pred.rf<-predict(rf,valid)
```
A look at the confusion matrix.
```{r echo=TRUE}
confusionMatrix(pred.rf,valid$classe)
```
We can also look at the variable importance obtained from random forest
```{r echo=TRUE}
rf.importance<-rf$importance
rf.importance
```

##Predict the test cases
Finally the model was applied to the test dataset to predict the test cases
```{r echo=TRUE}
pred.test<-predict(rf,test)
pred.test
```


